%/* ----------------------------------------------------------- */
%/*                                                             */
%/*                          ___                                */
%/*                       |_| | |_/   SPEECH                    */
%/*                       | | | | \   RECOGNITION               */
%/*                       =========   SOFTWARE                  */ 
%/*                                                             */
%/*                                                             */
%/* ----------------------------------------------------------- */
%/*         Copyright: Microsoft Corporation                    */
%/*          1995-2000 Redmond, Washington USA                  */
%/*                    http://www.microsoft.com                */
%/*                                                             */
%/*   Use of this software is governed by a License Agreement   */
%/*    ** See the file License for the Conditions of Use  **    */
%/*    **     This banner notice must not be removed      **    */
%/*                                                             */
%/* ----------------------------------------------------------- */
%
% HTKBook - Steve Young  23/09/95
%

% Define standard option descriptions

%\stdopts{toolname}
\newcommand{\stdopts}[1]{
\htool{#1} also supports the standard options \texttt{-A},
\texttt{-C},  \texttt{-D}, \texttt{-S}, \texttt{-T}, and \texttt{-V} as described
in section~\ref{s:stdopts}.
}

\newcommand{\stdoptB}{
  \ttitem{-B} Output HMM definition files in binary format.
}
\newcommand{\stdoptE}{
  \ttitem{-E dir [ext]} Parent transform directory and optional extension
	for parent transforms. The default option is that no parent
	transform is used.
}
\newcommand{\stdoptF}{
  \ttitem{-F fmt} Set the source data format to {\tt fmt}.
}
\newcommand{\stdoptG}{
  \ttitem{-G fmt} Set the label file format to {\tt fmt}.
}
\newcommand{\stdoptH}{
  \ttitem{-H mmf} Load HMM macro model file {\tt mmf}.  This option
     may be repeated to load multiple MMFs.
}
\newcommand{\stdoptI}{
  \ttitem{-I mlf} This loads the master label file {\tt mlf}.  This option
     may be repeated to load several MLFs.
}
\newcommand{\stdoptJ}{
  \ttitem{-J dir [ext]} Add directory to the list of possible input
	transform directories. Only one of the options can specify the
	extension to use for the input transforms.
}
\newcommand{\stdoptK}{
  \ttitem{-K dir [ext]} Output transform directory and optional extension
	for output transforms. The default option is that there is no
	output extension and the current transform directory is used.
}
\newcommand{\stdoptL}{
  \ttitem{-L dir} Search directory \texttt{dir} for label files (default
     is to search current directory).
}
\newcommand{\stdoptM}{
  \ttitem{-M dir} Store output HMM macro model files in the directory \texttt{dir}.  
     If this option is not given, the new HMM definition will overwrite
     the existing one.
}
\newcommand{\stdoptO}{
  \ttitem{-O fmt} Set the target data format to {\tt fmt}.
}
\newcommand{\stdoptP}{
  \ttitem{-P fmt} Set the target label format to {\tt fmt}.
}
\newcommand{\stdoptQ}{
  \ttitem{-Q} Print a summary of all commands supported by this tool.
}
\newcommand{\stdoptX}{
  \ttitem{-X ext} Set label file extension to {\tt ext} 
  (default is \texttt{lab}).
}


%%% Local Variables: 
%%% mode: plain-tex
%%% TeX-master: "../htkbook"
%%% End: 
